---
title: "VarSelLCM"
author: "Variable Selection for Model-Based Clustering of Mixed-Type Data Set with  Missing Values"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette VarSelLCM}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---



**References:** 

* Marbac, M. and Sedki, M. (2017), [Variable selection for model-based clustering using the integrated complete-data likelihood](https://link.springer.com/article/10.1007/s11222-016-9670-1), Statistics and Computing, Volume 27, Issue 4, pp 1049–1063.
* Marbac, M., Patin, E. and Sedki, M. (2018), [Variable selection for mixed data clustering: Application in human population genomics](https://arxiv.org/abs/1703.02293), Arxiv 1703.02293.

# Introduction
*[VarSelLCM](https://CRAN.R-project.org/package=VarSelLCM)* permits a full model selection (detection of the relevant features for clustering and selection of the number of clusters) in model-based clustering, according to classical information criteria (BIC, MICL or AIC).

Data to analyzed can be composed of continuous, integer and/or categorical features. Moreover, missing
values are managed, without any pre-processing, by the model used to cluster with the assumption that
values are missing completely at random. 
Thus, VarSelLCM can also be used for data imputation via mixture models. 

An R-Shiny application is implemented to easily interpret the clustering results


# Mixed-type data analysis

This section performs the whole analysis of the *Heart* data set. *Warning continuous features must be stored in numeric, integer features must be stored in integer and categorical features must be stored in factor.*
```{r, comment=""}
library(VarSelLCM)
# Data loading
data("heart")
head(heart)
```

Clustering is performed with variable selection. 
Model selection is done with BIC because n>>d.
The number of components is between 1 and 4.
Do not hesitate to use parallelisation (here only two cores are used).

```{r, comment=""}
# Add a missing value artificially (just to show that it works!)
heart[1,1] <- NA
# Clustering with variable selection and a number of cluster betwee 1 and 4
# Model selection is BIC (to use MICL, the option must be specified)
out <- VarSelCluster(heart[,-13], 1:4, nbcores = 2)
```

Now, all the results can be analyzed by the Shiny application...
```{r, eval=FALSE, comment="", include=TRUE}
# Start the shiny application
 VarSelShiny(out)
```

... but this analysis can also be done on R.

To get a summary of the selected model.
```{r, comment=""}
# Summary of the best model
summary(out)
```

Model interpretation should focus on the most discriminative variables. 
These variables can be found with the following plot.
```{r, comment=""}
# Discriminative power of the variables (here, the most discriminative variable is MaxHeartRate)
plot(out, type="bar")
```

Interpretation of the most discriminative variable is based on its distribution per cluster.
```{r, comment=""}
# Boxplot for continuous (or interger) variable
plot(out, y="MaxHeartRate", type="boxplot")
```

We can check that the distribution used to cluster is relevant.
```{r, comment=""}
# Empirical and theoretical distributions (to check that clustering is pertinent)
plot(out, y="MaxHeartRate", type="cdf")
```

Interpretation of a categorical variable is based on its distribution per cluster.
```{r, comment=""}
# Summary of categorical variable
plot(out, y="Sex")
```

Interpretation of the cluster overlaps by using the probabilities of missclassification.
```{r, comment=""}
# Summary of the probabilities of missclassification
plot(out, type="probs-class")
```

Missing values can be imputed.
```{r, comment=""}
# Imputation by posterior mean for the first observation
not.imputed <- heart[1,-13]
imputed <- VarSelImputation(out)[1,]
rbind(not.imputed, imputed)
```